Implementation of a lattice Boltzmann method for immiscible multiphase flow simulations using the level set method

### 2015-11-13 Paper/paper.tex
Korrekturen am Paper

\author
{Lorenz Hufnagel, Daniel Zint\\
Chair for System-Simulation, Friedrich Alexander Univerit\"at Erlangen-N\"urnberg\\
Cauerstr. 11, 91058 Erlangen, Germany\\
lorenz.hufnagel@studium.uni-erlangen.de\\
daniel.zint@fau.de % ist das so richtig?
}

@@

The two-phase flow problem can be described with the Navier-Stokes equations for two incompressible fluids. 
%TODO Regina: Hier fehlen Referenzen! ausserdem muessen noch die Gleichungen unten gelabelt werden!!!
To apply them, we distribute our domain into three parts: $\Omega_1$ and $\Omega_2$ describe the two fluids and $\Gamma = \delta\Omega_1 \cap \delta\Omega_2$ is the interface (Fig. \ref{intro_phases}). The material properties in each fluid are constant and the interface is assumed to be sharp. Thus the material properties are discontinuous and the fluids will not mix.

@@

The evolution of the flow fields within each fluid can be expressed with the incompressible Navier-Stokes equations,
\begin{equation}
	\nabla \cdot  \vec u_i = 0 \text{  in } \Omega_i \text{,}
\end{equation}
\begin{equation}
\frac{\partial \vec u_i}{\partial t} + \left(\vec  u_i\cdot \nabla \right) \vec u_i
=
-\frac{1}{\varrho_i}\nabla p_i + \nu_i \nabla^2 \vec u_i \text{  in } \Omega_i \text{.}
\end{equation}
Here, $u_i$ is the velocity, $p_i$ the pressure, $\nu_i$ the kinematic viscosity and $\varrho_i$ the mass density.

As boundary conditions and initial conditions we can take for example,
\begin{align}
	\vec u_i(x,t) = 0 \text{, on } \delta\Omega_i \backslash \Gamma \text{,} \\
	\vec u_i(x,0) = \vec u_{ini,i}(x) \text{, in } \Omega_i \text{.}
\end{align}
For the interface $\Gamma$ we have the influence of the surface tension and a no-slip condition, i.e. continuity of the velocities between the fluid phases. Therefore, we get the jump conditions,
\begin{align}
	[\vec u] = 0 \text{,  } \\
	[T] \cdot \vec n = 2 \sigma \kappa \vec n \text{.}
	\label{jumpconditions}
\end{align}
Here, T is the shear stress tensor, $\sigma$ the surface tension, and $\kappa$ the curvature of the interface with respect to its outer normal $\vec n$. Brackets denote the jump of a quantity at the interface, 
\begin{equation}
	[q](\vec x) = \lim_{\epsilon \rightarrow 0} (q(\vec x - \epsilon \vec n) - (q(\vec x + \epsilon \vec n)) \text{ with } \vec x \in \Gamma. % I also turned this around, according to [p].
\end{equation}
In Fig. \ref{intro_phases} this is the jump from $\Omega_1$ to $\Omega_2$. For the other direction the sign changes.

There are different approaches to solve the immiscible two-phase problem. The most spread among these approaches is the colour gradient method of Gunstensen and Rothman \cite{GRcg1,GRcg2,GRcg3}. The downside of this method is that it is only applicable for small density and viscosity differences. Another approach is the concept of {\it interaction potentials} by Shan and Chen \cite{ShanChen1,ShanChen2}. This method actually models miscible fluids and can only approximately describe the behavior of immiscible fluids. Furthermore there are free surface methods which only take one fluid in account for computation (\cite{Koerner}). They work best for big density and viscosity ratios.

@@

\subsection{The Lattice Boltzmann Method}
To solve the incompressible Navier-Stokes equations, we use the lattice Boltzmann method. This method describes the evolution of the particle density $f(\vec{x},\vec{v},t)$ in phase space, with $(\vec{x},\vec{v})$ as the phase space variables and $t$ as time. The LBM algorithm discretizes with a regular grid in space and with restricted number of velocities adapted to this grid \cite{LBM1,LBM2,LBM3,LBM4}. We use the D2Q9 model in 2D, which has 9 velocity vectors on a plane grid with unit spacing including one zero velocity (Fig. \ref{lbmc_i}). 
%TODO Regina: Man kann auch Cref benutzen zum Zitieren
\begin{figure}
	\hfill\includegraphics[width=6cm,natwidth=403,natheight=409]{LBMc_i.png}\hspace*{\fill}
	\caption{The vectors of one cell in LBM.}
	\label{lbmc_i}
\end{figure}
The vectors are given by the columns of the matrix,
\begin{equation}
	c = 
	\begin{bmatrix}
	0 & 0 & 1 & 1 &  1 &  0 & -1 & -1 & -1 \\
	0 & 1 & 1 & 0 & -1 & -1 & -1 &  0 &  1
	\end{bmatrix}
	\text{.}
\end{equation}
The corresponding particle distributions are denoted by $f_i(\vec{x},t) = f(\vec{x},\vec{c_i},t)$. We get the density and velocity from the particle distribution functions with,
\begin{equation}
	\rho(\vec{x},t) = \sum_{i=1}^{9} f_i(\vec{x},t) \text{, } \quad
	\vec{u} = \sum_{i=1}^{9} f_i(\vec{x},t)\vec{c_i} \text{.}
\end{equation}
For the collision step we use the Bhatnagar-Gross-Krook (BGK) approximation \cite{BGK}
\begin{equation}
	f_i(\vec{x}+\vec{c_i},t+1) = f_i(\vec{x},t) - \frac1\tau(f_i-f_i^{eq}) + G_i \text{.}
\end{equation}
The parameter $\tau$ is the relaxation parameter for the BGK collision operator and controls the kinematic viscosity in lattice units $\nu = \frac16(2\tau - 1)$. Note, that we assume the cell spacing $\Delta x$ and the time step $\Delta t$ to be unity here. The lattice speed of sound equals to $c_s =\frac{1}{\sqrt{3}}$. $G_i$ is an additional force, e.g. gravity. In our case we set $G_i = 0$. Furthermore, we use the equilibrium distribution,
\begin{equation}
	f_i^{eq}(\rho,\vec{u}) = f_i^*(\rho + 3\vec{c_i} \cdot \vec{u} + \frac92(\vec{c_i} \cdot \vec{u})^2 - \frac32 \vec u\cdot \vec u)
\end{equation}
with the corresponding D2Q9 weight factors,
$$
f_i^* = \begin{cases} \frac49 \text{,} & i = 0 \\ \frac19 \text{,} & i = 1,3,5,7 \\ \frac{1}{36} \text{,} &  i = 2,4,6,8 \end{cases} \text{.}
$$

One time step in the LBM algorithm is achieved by the following sub-steps:
\begin{itemize}
	\item[1. ] Collision step: $f_i^+ = f_i - \frac1\tau(f_i - f_i ^{eq}) + G_i$
	\item[2. ] Propagation step: $f_i(\vec{x}+\vec{c_i},t+1) = f_i^+(x,t)$, for interior nodes
	\item[3. ] Boundary conditions: $f_i(\vec{x}+\vec{c_i},t+1) = f_i(\vec{x}+\vec{c_i},t)$, if $\vec{x} \notin \Omega$
\end{itemize}
%TODO Regina: Propagation kurz im Text erklären! 
%TODO Regina: Bounce-Back boundary conditions erklaeren!

@@

Let $\Gamma_t$ be our interface at time t. Within the level set method, $\Gamma_t$ is the zero iso-surface of the level set function $\varphi$. Thus, for $\vec x \in \Gamma_t$ we have $\varphi(\vec x,t) = 0$, in direction of the surface normal we get $\varphi > 0$, and in the other direction we get $\varphi < 0$. For this level set function we get the level set equation introduced by Osher and Sethian in \cite{OsherSethian} for the evolution of free surfaces,
\begin{equation}
  \frac{\partial \varphi}{\partial t} + \vec v \cdot \nabla \varphi = 0 \text{,}
\end{equation}
with the velocity $\vec v = \vec u(\vec x(t))$ for $\vec x(t) \in \Gamma_t$, which we obtain from the lattice Boltzmann method.
%TODO Regina: LBM als Abkürzung einführen

@@

%TODO Regina: die Toolbox muss anders referenziert werden in der Bibliography. Als URL link oder eben die Homepage, wo man das runterladen kann und wann ihr es runtergeladen haben. Gibt dafuer MISC in bibtex

@@

For the coupling, we use a bounce back type boundary condition that prescribes the velocity as Dirichlet value. 
%TODO Regina: REF!!!
For example, in Fig. \ref{coupling} for a point $x_2 \in \Omega_2$ that is next to the phase interface we set,

@@

That is, an orthogonal projection of the column vectors of the (symmetric) deviatoric shear rate jump tensor on the link direction weighted with the interface distance and a compensation term for the interpolation error that arises, when the interface is not on the exact middle between two LBM grid cells.
%TODO Regina: sehr langer Satz, 2-3 einfache Saetze draus machen!
The tensor product is $a \otimes b = ab^T$ and the double contraction $A:B = \mathrm{trace}(AB^T)$. We can rewrite formula (\ref{add_term2}) in the form,

@@

\begin{align}

@@

\end{align}

@@

@@

\bibitem{Thoemmes} {\sc G. Th\"ommes, J. Becker}, {\em A lattice Boltzmann method for immiscible multiphase flow simulations using the level set method}, J. Comput. Phys. 228 (2009), 1139-1156

\bibitem{Adalsteinsson} {\sc D. Adalsteinsson, J.A. Sethian}, {\em A fast level set method for propagating interfaces}, J. Comput. Phys. 118 (2) (1995) 269-277

\bibitem{GRcg1} {\sc A.K. Gunstensen, D.H. Rothman, S. Zaleski, G. Zanetti}, {\em Lattice Boltzmann model of immiscible fluids}, Phys. Rev. A 43 (8) (1991) 4320-4327

\bibitem{GRcg2} {\sc A.K. Gunstensen, D.H. Rothman}, {\em Macroscopic modeling of immiscible flows in three dimensions by a lattice Boltzmann method}, Europhys. Lett. 18 (2) (1998) 157-161

\bibitem{GRcg3} {\sc K. Sankaranarayanan, I.G. Kevrekidies, S. Sundaresan, J. Lu, G. Tryggvason}, {\em A comparative study of lattice Boltzmann and front-tracking finite-difference methods for bubble simulations}, Int. J. Multiphase Flow 29 (1) (2003) 109-116

\bibitem{ShanChen1} {\sc X. Shan, H. Chen}, {\em Lattice Boltzmann model for simulation flows with multiple phases and components}, Phys. Rev. E 47 (3) (1999) 567-603

\bibitem{ShanChen2} {\sc X. Shan, H. Chen}, {\em Simulation of nonideal gases and liquid-gas phase transition by the Lattice Boltzmann equation}, Phys. Rev. E 49 (4) (1994) 2941

\bibitem{Koerner} {\sc C. K\"orner, M. Thies, T. Hofmann, N. Th\"urey, U. R\"ude}, {\em Lattice Boltzmann Model for Free Surface Flow for Modeling Foaming}, J. Stat. Phys. 121 (2005) 179-196

\bibitem{LBM1} {\sc S. Chen, G.D. Doolenm}, {\em Lattice Boltzmann method for fluid flows}, Ann. Rev. Fluid Mech. 30 (1998) 329-364

\bibitem{LBM2} {\sc X. He, L.S. Luo}, {\em Lattice Boltzmann model for the incompressible Navier-Stokes equation}, J. Stat. Phys. 88 (1997) 927-944

\bibitem{LBM3} {\sc S. Succi}, {\em The Lattice Boltzmann Equation for Fluid Dynamics and Beyond}, Clarendon Press, Oxford, ISBN 978-0-19-850398-9 (2001)

\bibitem{LBM4} {\sc D.A. Wolf-Gladow}, {\em Lattice-Gas Cellular Automata and Lattice Boltzmann Models}, An Introduction, Springer, 2000

\bibitem{BGK} {\sc P. Bhatnagar, E. Gross, M. Krook}, {\em A model for collision processes in gases: 1. Small amplitude processes in charged and neutral one-component systems}, Phys. Rev. 94 (1954) 511

\bibitem{OsherSethian} {\sc S. Osher, J. Sethian}, {\em Fronts propagating with curvature-dependent speed: algorithms based on Hamlton-Jacobi formulations}, J. Comput. Phys. 79 (1988) 12-49

\bibitem{mitchell}
Mitchell, Ian, A Toolbox of Level Set Methods

\bibitem{Caiazzo} {\sc A. Caiazzo}, {\em Asymptotic analysis of lattice Boltzmann method for fluid-structure interaction problems}, Ph.D. Thesis, Kaiserslautern (Germany) and Pisa (Italy), 2007

\bibitem{Krueger} {\sc T. Kr\"uger, F. Varnik, D. Raabe}, {\em Shear stress in lattice Boltzmann simulations}, Phys. Rev. E 79 (4) (2009) 046704

\bibitem{Lallemand} {\sc P. Lallemand, L.-S. Luo}, {\em Lattice Boltzmann method for moving boundaries}, J. Comput. Phys. 184 (2003) 406-421

>>>>

\author
{Lorenz Hufnagel, Daniel Zint\\
Chair for System-Simulation, Friedrich Alexander Univerit\"at Erlangen-N\"urnberg\\
Cauerstr. 11, 91058 Erlangen, Germany\\
lorenz.hufnagel@fau.de\\
daniel.zint@fau.de % ist das so richtig?
}

@@

The two-phase flow problem can be described with the Navier-Stokes equations for two incompressible fluids (Eq. \ref{nse1}, \ref{nse2}). 
%TODO Regina: Hier fehlen Referenzen!
To apply them, we distribute our domain into three parts: $\Omega_1$ and $\Omega_2$ describe the two fluids and $\Gamma = \delta\Omega_1 \cap \delta\Omega_2$ is the interface (Fig. \ref{intro_phases}). The material properties in each fluid are constant and the interface is assumed to be sharp. Thus the material properties are discontinuous and the fluids will not mix.

@@

The evolution of the flow fields within each fluid can be expressed with the incompressible Navier-Stokes equations,
\begin{gather}
	\nabla \cdot  \vec u_i = 0 \text{  in } \Omega_i \text{,}
	\label{nse1} \\
	\frac{\partial \vec u_i}{\partial t} + \left(\vec  u_i\cdot \nabla \right) \vec u_i
	=
	-\frac{1}{\varrho_i}\nabla p_i + \nu_i \nabla^2 \vec u_i \text{  in } \Omega_i \text{.}
	\label{nse2}
\end{gather}
Here, $u_i$ is the velocity, $p_i$ the pressure, $\nu_i$ the kinematic viscosity and $\varrho_i$ the mass density.

As boundary conditions and initial conditions we can take for example,
\begin{gather}
	\vec u_i(x,t) = 0 \text{, on } \delta\Omega_i \backslash \Gamma \text{,} \\
	\vec u_i(x,0) = \vec u_{ini,i}(x) \text{, in } \Omega_i \text{.}
\end{gather}
For the interface $\Gamma$ we have the influence of the surface tension and a no-slip condition, i.e. continuity of the velocities between the fluid phases. Therefore, we get the jump conditions,
\begin{gather}
	[\vec u] = 0 \text{,  } \\
	[T] \cdot \vec n = 2 \sigma \kappa \vec n \text{.}
	\label{jumpconditions}
\end{gather}
Here, T is the shear stress tensor, $\sigma$ the surface tension, and $\kappa$ the curvature of the interface with respect to its outer normal $\vec n$. Brackets denote the jump of a quantity at the interface, 
\begin{equation}
	[q](\vec x) = \lim_{\epsilon \rightarrow 0} (q(\vec x - \epsilon \vec n) - (q(\vec x + \epsilon \vec n)) \text{ with } \vec x \in \Gamma. % I also turned this around, according to [p].
\end{equation}
In Fig. \ref{intro_phases} this is the jump from $\Omega_1$ to $\Omega_2$. For the other direction the sign changes.

There are different approaches to solve the immiscible two-phase problem. The most spread among these approaches is the colour gradient method of Gunstensen and Rothman \cite{GRcg1,GRcg2,GRcg3}. The downside of this method is that it is only applicable for small density and viscosity differences. Another approach is the concept of {\it interaction potentials} by Shan and Chen \cite{ShanChen1,ShanChen2}. This method actually models miscible fluids and can only approximately describe the behavior of immiscible fluids. Furthermore there are free surface methods which only take one fluid in account for computation (\cite{Koerner}). They work best for big density and viscosity ratios.
%TODO: Mehr Referenzen, siehe Mail

@@

\subsection{The Lattice Boltzmann Method}
To solve the incompressible Navier-Stokes equations, we use the lattice Boltzmann method (LBM). This method describes the evolution of the particle density $f(\vec{x},\vec{v},t)$ in phase space, with $(\vec{x},\vec{v})$ as the phase space variables and $t$ as time. The LBM algorithm discretizes with a regular grid in space and with restricted number of velocities adapted to this grid \cite{LBM1,LBM2,LBM3,LBM4}. We use the D2Q9 model in 2D, which has 9 velocity vectors on a plane grid with unit spacing including one zero velocity (Fig. \ref{lbmc_i}). 
%TODO Regina: Man kann auch Cref benutzen zum Zitieren
\begin{figure}
	\hfill\includegraphics[width=6cm,natwidth=403,natheight=409]{LBMc_i.png}\hspace*{\fill}
	\caption{The vectors of one cell in LBM.}
	\label{lbmc_i}
\end{figure}
The vectors are given by the columns of the matrix,
\begin{equation}
	c = 
	\begin{bmatrix}
	0 & 0 & 1 & 1 &  1 &  0 & -1 & -1 & -1 \\
	0 & 1 & 1 & 0 & -1 & -1 & -1 &  0 &  1
	\end{bmatrix}
	\text{.}
\end{equation}
The corresponding particle distributions are denoted by $f_i(\vec{x},t) = f(\vec{x},\vec{c_i},t)$. We get the density and velocity from the particle distribution functions with,
\begin{equation}
	\rho(\vec{x},t) = \sum_{i=1}^{9} f_i(\vec{x},t) \text{, } \quad
	\vec{u} = \sum_{i=1}^{9} f_i(\vec{x},t)\vec{c_i} \text{.}
\end{equation}

In the interior, the LBM alternates collision step and propagation step. For the collision step we use the Bhatnagar-Gross-Krook (BGK) approximation \cite{BGK},
\begin{equation}
	f_i^+(\vec{x}+\vec{c_i},t+1) = f_i(\vec{x},t) - \frac1\tau(f_i-f_i^{eq}) + G_i \text{.}
\end{equation}
The propagation step is denoted by,
\begin{equation}
	f_i(\vec{x}+\vec{c_i},t+1) = f_i^+(\vec{x},t+1).
\end{equation}
The parameter $\tau$ is the relaxation parameter for the BGK collision operator and controls the kinematic viscosity in lattice units $\nu = \frac16(2\tau - 1)$. Note, that we assume the cell spacing $\Delta x$ and the time step $\Delta t$ to be unity here. The lattice speed of sound equals to $c_s =\frac{1}{\sqrt{3}}$. $G_i$ is an additional force, e.g. gravity. In our case we set $G_i = 0$. Furthermore, we use the equilibrium distribution,
\begin{equation}
	f_i^{eq}(\rho,\vec{u}) = f_i^*(\rho + 3\vec{c_i} \cdot \vec{u} + \frac92(\vec{c_i} \cdot \vec{u})^2 - \frac32 \vec u\cdot \vec u)
\end{equation}
with the corresponding D2Q9 weight factors,
$$
f_i^* = \begin{cases} \frac49 \text{,} & i = 0 \\ \frac19 \text{,} & i = 1,3,5,7 \\ \frac{1}{36} \text{,} &  i = 2,4,6,8 \end{cases} \text{.}
$$

%Boundary Conditions: Bounce-back & moving no-slip & inflow-outflow
We use two types of boundary conditions. To simulate static walls, we use simple bounce-back boundary conditions,
\begin{equation}
	\tilde{f_i} = f_{i^*}^+(\vec{x_b},t) \text{, with } \vec{x_b} \in \delta\Omega \text{,}
\end{equation}
where $i^*$ is the index of the opposite direction $c_{i^*} = -c_i$. To simulate a flow, we use moving no-slip boundary conditions,
\begin{equation}
	\tilde{f_i} = f_{i^*}^+(\vec{x_b},t) + 6 f_i^*\vec{c_i} \vec{v} \text{, with } \vec{x_b} \in \delta\Omega,
\end{equation}
where $\vec{v}$ is the speed of the moving wall. To enable inflow and outflow of fluids we have periodic boundary conditions,
\begin{equation}
	\tilde{f_i}(\vec{x_a},t) = f_i^+(\vec{x_b},t) \text{,}
\end{equation}
where $x_a$ and $x_b$ are points of opposite boundaries.

One time step in the LBM algorithm is achieved by the following sub-steps:
\begin{itemize}
	\item[1. ] Collision step: $f_i^+ = f_i - \frac1\tau(f_i - f_i ^{eq}) + G_i$
	\item[2. ] Propagation step: $f_i(\vec{x}+\vec{c_i},t+1) = f_i^+(x,t)$, for interior nodes
	\item[3. ] Boundary conditions: $f_i(\vec{x}+\vec{c_i},t+1) = f_i(\vec{x}+\vec{c_i},t)$, if $\vec{x} \notin \Omega$
\end{itemize}

@@

Let $\Gamma_t$ be our interface at time t. Within the level set method, $\Gamma_t$ is the zero iso-surface of the level set function $\varphi$. Thus, for $\vec x \in \Gamma_t$ we have $\varphi(\vec x,t) = 0$, in direction of the surface normal we get $\varphi > 0$, and in the other direction we get $\varphi < 0$. For this level set function we get the level set equation introduced by Osher and Sethian in \cite{OsherSethian} for the evolution of free surfaces,
\begin{equation}
  \frac{\partial \varphi}{\partial t} + \vec v \cdot \nabla \varphi = 0 \text{,}
\end{equation}
with the velocity $\vec v = \vec u(\vec x(t))$ for $\vec x(t) \in \Gamma_t$, which we obtain from the LBM.

@@

@@

For the coupling, we use a bounce back type boundary condition that prescribes the velocity as Dirichlet value as described in \cite[p. 1146]{Thoemmes}. For example, in Fig. \ref{coupling} for a point $x_2 \in \Omega_2$ that is next to the phase interface we set,

@@

That is an orthogonal projection of the column vectors of the (symmetric) deviatoric shear rate jump tensor on the link direction weighted with the interface distance and a compensation term for the interpolation error. This error arises, when the interface is not on the exact middle between two LBM grid cells. The tensor product is $a \otimes b = ab^T$ and the double contraction $A:B = \mathrm{trace}(AB^T)$. We can rewrite formula (\ref{add_term2}) in the form,

@@

\begin{gather}

@@

\end{gather}

@@

%TODO: Kommentar, dass hier ein Fehler im Thömmes-paper ist

@@

\bibitem{Thoemmes} {\sc G. TH\"OMMES, J. BECKER}, A lattice Boltzmann method for immiscible multiphase flow simulations using the level set method, J. Comput. Phys. 228 (2009), 1139-1156.

\bibitem{Adalsteinsson} {\sc D. ADALSTEINSSON, J.A. SETHIAN}, A fast level set method for propagating interfaces, J. Comput. Phys. 118 (2) (1995) 269-277.

\bibitem{GRcg1} {\sc A.K. GUNSTENSEN, D.H. ROTHMAN, S. ZALESKI, G. ZANETTI}, Lattice Boltzmann model of immiscible fluids, Phys. Rev. A 43 (8) (1991) 4320-4327.

\bibitem{GRcg2} {\sc A.K. GUNSTENSEN, D.H. ROTHMAN}, Macroscopic modeling of immiscible flows in three dimensions by a lattice Boltzmann method, Europhys. Lett. 18 (2) (1998) 157-161.

\bibitem{GRcg3} {\sc K. SANKARANARAYANAN, I.G. KEVREKIDIES, S. SUNDARESAN, J. LU, G. TRYGGVASON}, A comparative study of lattice Boltzmann and front-tracking finite-difference methods for bubble simulations, Int. J. Multiphase Flow 29 (1) (2003) 109-116.

\bibitem{ShanChen1} {\sc X. SHAN, H. CHEN}, Lattice Boltzmann model for simulation flows with multiple phases and components, Phys. Rev. E 47 (3) (1999) 567-603.

\bibitem{ShanChen2} {\sc X. SHAN, H. CHEN}, Simulation of nonideal gases and liquid-gas phase transition by the Lattice Boltzmann equation, Phys. Rev. E 49 (4) (1994) 2941.

\bibitem{Koerner} {\sc C. K\"ORNER, M. THIES, T. HOFMANN, N. TH\"UREY, U. R\"UDE}, Lattice Boltzmann Model for Free Surface Flow for Modeling Foaming, J. Stat. Phys. 121 (2005) 179-196.

\bibitem{LBM1} {\sc S. CHEN, G.D. DOOLENM}, Lattice Boltzmann method for fluid flows, Ann. Rev. Fluid Mech. 30 (1998) 329-364.

\bibitem{LBM2} {\sc X. HE, L.S. LUO}, Lattice Boltzmann model for the incompressible Navier-Stokes equation, J. Stat. Phys. 88 (1997) 927-944.

\bibitem{LBM3} {\sc S. SUCCI}, The Lattice Boltzmann Equation for Fluid Dynamics and Beyond, Clarendon Press, Oxford, ISBN 978-0-19-850398-9 (2001).

\bibitem{LBM4} {\sc D.A. WOLF-GLADOW}, Lattice-Gas Cellular Automata and Lattice Boltzmann Models, An Introduction, Springer, 2000.

\bibitem{BGK} {\sc P. BHATNAGAR, E. GROSS, M. KROOK}, A model for collision processes in gases: 1. Small amplitude processes in charged and neutral one-component systems, Phys. Rev. 94 (1954) 511.

\bibitem{OsherSethian} {\sc S. OSHER, J. SETHIAN}, Fronts propagating with curvature-dependent speed: algorithms based on Hamlton-Jacobi formulations, J. Comput. Phys. 79 (1988) 12-49.

\bibitem{mitchell}I. MITCHELL, A Toolbox of Level Set Methods, August 2015, URL: \url{http://www.cs.ubc.ca/~mitchell/ToolboxLS/}.

\bibitem{Caiazzo} {\sc A. CAIAZZO}, Asymptotic analysis of lattice Boltzmann method for fluid-structure interaction problems, Ph.D. Thesis, Kaiserslautern (Germany) and Pisa (Italy), 2007.

\bibitem{Krueger} {\sc T. KR\"UGER, F. VARNIK, D. RAABE}, Shear stress in lattice Boltzmann simulations, Phys. Rev. E 79 (4) (2009) 046704.

\bibitem{Lallemand} {\sc P. LALLEMAND, L.-S. LUO}, Lattice Boltzmann method for moving boundaries, J. Comput. Phys. 184 (2003) 406-421.
